\textcolor{blue}{A gradient-based optimization method for motion planning}$^{1}$

### 2016-01-28 paper.tex
Fix duplicated parts in 3.6 and typo

In the collision-configuration $\xx_{i+1}(\tcolli)$, let $\po_c\in \real^3$ be a 
contact point expressed in the global frame 
(Figure~\ref{contact-points} left).
\textcolor{blue}{We denote by
  \begin{itemize}
  \item $\po_1^{loc}$ (resp. $\po_2^{loc}$) the coordinate vector of $\po_c$ in the local frame of  $\body_1$ (resp. $\body_2$),
  \item $M_1(\conf) \in SE(3)$ (resp. $M_2(\conf) \in SE(3)$) the rigid-body transformation representing the position of $\body_1$ (resp. $\body_2$) in the global frame, in configuration $\conf$,
  \item $M_2^1 (\conf) = M_1(\conf)^{-1} M_2(\conf)$ the position of $\body_2$ local frame in $\body_1$ local frame in configuration $\conf$,
  \item $\po_1(\conf)$ (resp. $\po_2(\conf)$) the points moving with $\body_1$ (resp. $\body_2$) of local coordinate $\po_1^{loc}$ (resp. $\po_2^{loc}$) in $\body_1$ (resp. $\body_2$) local frame.
  %% \item $\rotation_1(\conf) \in SO(3)$ (resp. $\rotation_2(\conf) \in SO(3)$) the rotation associated with $M_1(\conf)$ (resp. $M_2(\conf)$).
  \end{itemize}
We define $\U$ as the coordinate vector of the unit vector linking points $\po_1$ and $\po_2$ in configuration $\xx_{i}(\tcolli)$, expressed in local frame of $\body_1$ (Figure~\ref{contact-points} right):
$$
\U = \frac{M_2^1 (\xx_{i}(\tcolli))\po_2^{loc} - \po_1^{loc}}{\|M_2^1 (\xx_{i}(\tcolli))\po_2^{loc} - \po_1^{loc}\|}
$$
Note that $\U$ is well defined since configuration $\xx_{i}(\tcolli)$ is collision-free.
}
\textcolor{blue}{
  Let $g$ be the real valued function mapping the projection of vector $\overrightarrow{\po_1\,\po_2(\conf)}$ on $\mathbf{u}$ to a configuration $\conf$ . For any $\conf \in \CS$:
\begin{equation}\label{eq:g}
g (\conf) = \left(M_2^1 (\conf)\po_2^{loc} - \po_1^{loc} | \U\right)
\end{equation}
}
%which is calculated from the last collision-free path at abscissa $\tcolli$.

@@

\textcolor{blue}{
Therefore the following expressions
\begin{align*}
\alpha_i \frac{\partial f}{\partial \xx} (\xx_i)\p_i = \frac{d}{dt}g (\traj (0)) & &
\alpha_i \frac{\partial f}{\partial \xx} (\xx_{i+1})\p_i =\frac{d}{dt}g (\traj (1))
\end{align*}
correspond to $(\velocity_{\po_2/\body_1}|\mathbf{u})$, respectively in configurations $\xx_{i}(\tcolli)$ and $\xx_{i+1}(\tcolli)$, where $\velocity_{\po_2/\body_1}$ represents the velocity of point $P_2$ in reference frame $\body_1$. 
Note that Assumption~(\ref{eq:assumption-1}) is equivalent to the first above equality to be different from 0. According to the geometric interpretation, Assumption~(\ref{eq:assumption-1}) is violated if and only if $\velocity_{\po_2/\body_1}$ is orthogonal to $\U$. Although very unlikely, this case might appear for some new constraint. In this case, inserting constraint $L_{j+1}$ is useless since
$$
\kernel \Jf_i = \kernel \Jf_{i+1}.
$$
}

@@

\textcolor{blue}{
When the new constraint is not linearly independent from the set of previous constraints, the algorithm enters an additional loop, performed by Algorithm~\ref{algo:solveRedundant}, in order to find a new constraint that is linearly independent. The loop keeps looking for paths along line segment $[\xx_{i},\xx_{i+1}]$ by dichotomy. A pair containing the latest free path and the latest path in collision, denoted by $(\xx_{Free}, \xx_{Coll})$ is stored along the loop. New iterations are chosen in the middle of this pair. 
\begin{itemize}
\item If the new path is collision-free, it replaces $\xx_{Free}$ in the pair.
\item If the new path is in collision, it replaces $\xx_{Coll}$ in the pair.
\end{itemize}
In both cases, a new constraint is built following the method described in Section~\ref{subsubsec:new-constraint}. Two cases are then possible:
\begin{enumerate}
\item at some point in the loop the new constraint is linearly independent from the previous constraints. The new constraint is added to $\Jf_{i}$ to give rise to $\Jf_{i+1}$ and the loop is interrupted, or
\item each new constraint is linearly dependent from the previous constraints and the loop never ends.
\end{enumerate}
In the second case, the iterations of the loop converge to a path that we denote by $\bar{\xx}$. $\xx_{Free}$ and $\xx_{Coll}$ also both converge to $\bar{\xx}$. Note that $\bar{\xx}$ lies at the boundary between free paths and paths in collision along the search line. $\bar{\xx}$ necessarily lies at the boundary between free paths and paths in collision. Let us denote by $\bar{\kappa}$ the abscissa along $\bar{\xx}$ where $\body_1$ and $\body_2$ come to contact, and let us denote by $\bar{\po}$ the contact point.
}

@@

\textcolor{blue}{
\begin{property}\label{prop:convergence}
As long as along iteration $\p_i$, the trajectory $(\xx_i+t\alpha_i\p_i)(\bar{\kappa})$ does not enter in collision with contact point velocity equal to 0 in the frame of $\body_1$, Algorithm~\ref{algo:solveRedundant} converges in a finite number of steps.
\end{property}
The above property means that the gradient-based path length optimization algorithm described in this paper converges in all cases, except for ill-defined problems.
}

\textcolor{blue}{
Therefore the following expressions
\begin{align*}
\alpha_i \frac{\partial f}{\partial \xx} (\xx_i)\p_i = \frac{d}{dt}g (\traj (0)) & &
\alpha_i \frac{\partial f}{\partial \xx} (\xx_{i+1})\p_i =\frac{d}{dt}g (\traj (1))
\end{align*}
correspond to $(\velocity_{\po_2/\body_1}|\mathbf{u})$, respectively in configurations $\xx_{i}(\tcolli)$ and $\xx_{i+1}(\tcolli)$, where $\velocity_{\po_2/\body_1}$ represents the velocity of point $\po_2$ in reference frame $\body_1$. 
Note that Assumption~(\ref{eq:assumption-1}) is equivalent to the first above equality to be different from 0. According to the geometric interpretation, Assumption~(\ref{eq:assumption-1}) is violated if and only if $\velocity_{\po_2/\body_1}$ is orthogonal to $\U$. Although very unlikely, this case might appear for some new constraint. In this case, inserting constraint $L_{j+1}$ is useless since
$$
\kernel \Jf_i = \kernel \Jf_{i+1}.
$$
}

\textcolor{blue}{
When the new constraint is not linearly independent from the set of previous constraints, the algorithm enters an additional loop, performed by Algorithm~\ref{algo:solveRedundant}, in order to find a new constraint that is linearly independent. The loop keeps looking for paths along line segment $[\xx_{i},\xx_{i+1}]$ by dichotomy. A pair containing the latest free path and the latest path in collision, denoted by $(\xx_{Free}, \xx_{Coll})$ is stored along the loop. New iterations are chosen in the middle of this pair. 
\begin{itemize}
\item If the new path is collision-free, it replaces $\xx_{Free}$ in the pair.
\item If the new path is in collision, it replaces $\xx_{Coll}$ in the pair.
\end{itemize}
In both cases, a new constraint is built following the method described in Section~\ref{subsubsec:new-constraint}. Two cases are then possible:
\begin{enumerate}
\item at some point in the loop the new constraint is linearly independent from the previous constraints. The new constraint is added to $\Jf_{i}$ to give rise to $\Jf_{i+1}$ and the loop is interrupted, or
\item each new constraint is linearly dependent from the previous constraints and the loop never ends.
\end{enumerate}
In the second case, the iterations of the loop converge to a path that we denote by $\bar{\xx}$. $\xx_{Free}$ and $\xx_{Coll}$ also both converge to $\bar{\xx}$. Note that $\bar{\xx}$ lies at the boundary between free paths and paths in collision along the search line.
}

Doing so, either the new constraint makes the dimension of the kernel decrease, or 
makes the configuration in collision ($\xx_{i+1}$ or $\xx_{i+2}$ depending on the 
case above) and the latest collision-free path ($\xx_{i}$ or $\xx_{i+2}$) become 
closer and closer. And the continuity of $f$ implies that $\dot{\conf}(0)$ and 
$\dot{\conf}(1)$
converge toward the norm of the velocity of point $\po_2$ in the reference frame of 
$\body_1$ along trajectory~(\ref{eq:trajectory}), when $\po_2$ and $\po_1$ come to 
contact. We already handled this situation previously, admitting that the velocity norm 
would not vanish at $t=1$. Therefore, assumption~(\ref{eq:assumption-1}) is verified.
%$$
%\frac{\partial f}{\partial \xx}(\xx_i)\p_i \ \ \ \mbox{and}\ \ \ \frac{\partial f}{\partial \xx}(\xx_{i+1})\p_i
%$$

@@

\begin{algorithm}
\begin{algorithmic}%[1] % line numerotation
\INPUT{path to optimize $\xx_0$}
\OUTPUT{optimized collision-free path $\xx_0$}
\State $\alpha \gets \alpha_{init}$
\State $minReached \gets false$
\While{($\textbf{not}(noCollision \textbf{ and } minReached)$)}
	\State $\p = \texttt{computeIterate}()$
	\State $minReached = (||\p||<10^{-3} \textbf{ or } \alpha=1)$
	\State $\xx_1 \gets \xx_0 + \alpha\,\p$
	\If{(\textbf{not}(\texttt{validatePath($\xx_1$)}))}
		\State $noCollision \gets false$
		\If{$(\alpha \neq 1)$}
			\State \texttt{computeCollisionConstraint}($\xx_1$, $\xx_0$)
			\State \texttt{findNewConstraint}()
			\State \texttt{addCollisionConstraint}()
			\State $\alpha \gets 1$
		\Else
			\State $\alpha \gets \alpha_{init}$
		\EndIf
	\Else
		\State $\xx_0 \gets \xx_1$
		\State $noCollision \gets true$
	\EndIf
\EndWhile
\end{algorithmic}
\caption{Gradient-based (GB) algorithm for path-optimization.} \label{algo:gradient}
\end{algorithm}

@@

@@

\begin{figure}
	\centering
	\includegraphics[width=7.8cm]{contact_points6.pdf}
	\includegraphics[width=8cm]{contact_points2potential.pdf}
	\caption{Path-optimization results on 2D robots, moving around 
	gray obstacles. Initial paths are dashed and crosses represent contact points 
	$\po_c$ related to collision-constraints. Note that, on the left, the detour 
	completely disappears.}
	\label{2D_long}
\end{figure}

>>>>

In the collision-configuration $\xx_{i+1}(\tcolli)$, let $\po_c\in \real^3$ be a 
contact point expressed in the global frame 
(Figure~\ref{contact-points} left).
\textcolor{blue}{We denote by:
  \begin{itemize}
  \item $\po_1^{loc}$ (resp. $\po_2^{loc}$) the coordinate vector of $\po_c$ in the local frame of  $\body_1$ (resp. $\body_2$),
  \item $M_1(\conf) \in SE(3)$ (resp. $M_2(\conf) \in SE(3)$) the rigid-body transformation representing the position of $\body_1$ (resp. $\body_2$) in the global frame, in configuration $\conf$,
  \item $M_2^1 (\conf) = M_1(\conf)^{-1} M_2(\conf)$ the position of $\body_2$ local frame in $\body_1$ local frame in configuration $\conf$,
  \item $\po_1(\conf)$ (resp. $\po_2(\conf)$) the points moving with $\body_1$ (resp. $\body_2$) of local coordinate $\po_1^{loc}$ (resp. $\po_2^{loc}$) in $\body_1$ (resp. $\body_2$) local frame.
  %% \item $\rotation_1(\conf) \in SO(3)$ (resp. $\rotation_2(\conf) \in SO(3)$) the rotation associated with $M_1(\conf)$ (resp. $M_2(\conf)$).
  \end{itemize}
We define $\U$ as the coordinate vector of the unit vector linking points $\po_1$ and $\po_2$ in configuration $\xx_{i}(\tcolli)$, expressed in local frame of $\body_1$ (Figure~\ref{contact-points} right):
$$
\U = \frac{M_2^1 (\xx_{i}(\tcolli))\po_2^{loc} - \po_1^{loc}}{\|M_2^1 (\xx_{i}(\tcolli))\po_2^{loc} - \po_1^{loc}\|}
$$
Note that $\U$ is well defined since configuration $\xx_{i}(\tcolli)$ is collision-free.
}
\textcolor{blue}{
  Let $g$ be the real valued function mapping the projection of vector $\overrightarrow{\po_1\,\po_2(\conf)}$ on $\mathbf{u}$ to a configuration $\conf$ . For any $\conf \in \CS$:
\begin{equation}\label{eq:g}
g (\conf) = \left(M_2^1 (\conf)\po_2^{loc} - \po_1^{loc} | \U\right)
\end{equation}
}
%which is calculated from the last collision-free path at abscissa $\tcolli$.

@@

\textcolor{blue}{
Therefore the following expressions
\begin{align*}
\alpha_i \frac{\partial f}{\partial \xx} (\xx_i)\p_i = \frac{d}{dt}g (\traj (0)) & &
\alpha_i \frac{\partial f}{\partial \xx} (\xx_{i+1})\p_i =\frac{d}{dt}g (\traj (1))
\end{align*}
correspond to $(\velocity_{\po_2/\body_1}|\mathbf{u})$, respectively in configurations $\xx_{i}(\tcolli)$ and $\xx_{i+1}(\tcolli)$, where $\velocity_{\po_2/\body_1}$ represents the velocity of point $\po_2$ in reference frame $\body_1$. 
Note that Assumption~(\ref{eq:assumption-1}) is equivalent to the first above equality to be different from 0. According to the geometric interpretation, Assumption~(\ref{eq:assumption-1}) is violated if and only if $\velocity_{\po_2/\body_1}$ is orthogonal to $\U$. Although very unlikely, this case might appear for some new constraint. In this case, inserting constraint $L_{j+1}$ is useless since
$$
\kernel \Jf_i = \kernel \Jf_{i+1}
$$
}

@@

\textcolor{blue}{
When the new constraint is not linearly independent from the set of previous constraints, the algorithm enters an additional loop, performed by Algorithm~\ref{algo:solveRedundant}, in order to find a new constraint that is linearly independent. The loop keeps looking for paths along line segment $[\xx_{i},\xx_{i+1}]$ by dichotomy. A pair containing the latest free path and the latest path in collision, denoted by $(\xx_{Free}, \xx_{Coll})$ is stored along the loop. New iterations are chosen in the middle of this pair. 
\begin{itemize}
\item If the new path is collision-free, it replaces $\xx_{Free}$ in the pair.
\item If the new path is in collision, it replaces $\xx_{Coll}$ in the pair.
\end{itemize}
In both cases, a new constraint is built following the method described in Section~\ref{subsubsec:new-constraint}. Two cases are then possible:
\begin{enumerate}
\item at some point in the loop the new constraint is linearly independent from the previous constraints. The new constraint is added to $\Jf_{i}$ to give rise to $\Jf_{i+1}$, and the loop is interrupted, or
%the new constraint makes the dimension of the kernel decrease
\item each new constraint is linearly dependent from the previous constraints and the loop never ends.
\end{enumerate}
In the second case, the iterations of the loop converge to a path that we denote by $\bar{\xx}$. $\xx_{Free}$ and $\xx_{Coll}$ also both converge to $\bar{\xx}$. $\bar{\xx}$ necessarily lies at the boundary between free paths and paths in collision. Let us denote by $\bar{\kappa}$ the abscissa along $\bar{\xx}$ where $\body_1$ and $\body_2$ come to contact, and let us denote by $\bar{\po}$ the contact point.
}

@@

\textcolor{blue}{
\begin{property}\label{prop:convergence}
As long as along iteration $\p_i$, the trajectory $(\xx_i+t\alpha_i\p_i)(\bar{\kappa})$ does not enter in collision with contact point velocity equal to 0 in the frame of $\body_1$, Algorithm~\ref{algo:solveRedundant} converges in a finite number of steps.
\end{property}
The above property means that the gradient-based algorithm described in this paper converges in all cases, except for ill-defined problems.
}

@@

\begin{algorithm}[b]
\begin{algorithmic}%[1] % line numerotation
\INPUT{path to optimize $\xx_0$}
\OUTPUT{optimized collision-free path $\xx_0$}
\State $\alpha \gets \alpha_{init}$
\State $minReached \gets false$
\While{($\textbf{not}(noCollision \textbf{ and } minReached)$)}
	\State $\p = \texttt{computeIterate}()$
	\State $minReached = (||\p||<10^{-3} \textbf{ or } \alpha=1)$
	\State $\xx_1 \gets \xx_0 + \alpha\,\p$
	\If{(\textbf{not}(\texttt{validatePath($\xx_1$)}))}
		\State $noCollision \gets false$
		\If{$(\alpha \neq 1)$}
			\State \texttt{computeCollisionConstraint}($\xx_1$, $\xx_0$)
			\State \texttt{findNewConstraint}()
			\State \texttt{addCollisionConstraint}()
			\State $\alpha \gets 1$
		\Else
			\State $\alpha \gets \alpha_{init}$
		\EndIf
	\Else
		\State $\xx_0 \gets \xx_1$
		\State $noCollision \gets true$
	\EndIf
\EndWhile
\end{algorithmic}
\caption{Gradient-based (GB) algorithm for path-optimization.} \label{algo:gradient}
\end{algorithm}

@@

\begin{figure}[b]
	\centering
	\includegraphics[width=7.8cm]{contact_points6.pdf}
	\includegraphics[width=8cm]{contact_points2potential.pdf}
	\caption{Path-optimization results on 2D robots, moving around 
	gray obstacles. Initial paths are dashed and crosses represent contact points 
	$\po_c$ related to collision-constraints. Note that, on the left, the detour 
	completely disappears.}
	\label{2D_long}
\end{figure}

@@
